---
title: "Analysis JG For Meeting #2"
author: "Sin City Seeders"
date: "October 16, 2016"
output: 
  html_notebook: default
---



```{r setup, include=FALSE}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(dplyr)
library(DT)
library(readr)
library(readxl)
library(plotly)

```


```{r, include=FALSE}
experiment.data <- read_excel(path = '../data/EXPERIMENT DATA.xlsx', 
                              sheet = "TRAINING_DATA", col_names = TRUE)

```

<center><h3>Head of dataset </h3></center>
<BR>

```{r, echo= FALSE}
datatable(data = head(experiment.data, n = 1000))
```

<center><h3>Analysis</h3></center>
<br>

#### <span style="color:red"> Question: If a VARIETY is rejected in some stage, can it be used for a future experiment?
</span>

How do we know if a variety was rejected in some stage? Need a signal in the data (Kostas)

#### <span style="color:green">Alternatively, "Gap in Years" analysis </span>

For now, try to find a **gap in years** a variety's experiment and dig into that.

Method:

1. For each variety, find the YEARS that has experiments. (Eg. v000016: 2009, 2010, 2011)
2. Discover if there are any gaps between the initial and final years (Eg. V000016: GAP = 0)
3. If there are gaps, GAP = 1. Record in database.

Output:


```{r, echo= FALSE}
## Calculate statistics at the variety level

variety.years <- na.omit(unique(experiment.data[c("YEAR", "VARIETY", "CLASS_OF")]))
varieties <- na.omit(unique(experiment.data[c("VARIETY")]))
varieties <- varieties[order(varieties$VARIETY),]


for(variety.temp.0 in varieties$VARIETY){
  variety.temp <- as.character(variety.temp.0)
  variety.years.temp <- subset(variety.years, VARIETY == variety.temp)
  rYEARS.temp <- range(variety.years.temp$YEAR)
  rnYEARS.temp <- seq(rYEARS.temp[1], rYEARS.temp[2])
  gap.temp <- rnYEARS.temp[!rnYEARS.temp %in% variety.years.temp$YEAR]
#  gap.temp <- c(2011)
  if(length(gap.temp) == 0){
    varieties[varieties$VARIETY == variety.temp,2] <- 0
  } else {
    varieties[varieties$VARIETY == variety.temp,2] <- 1
  }
  varieties[varieties$VARIETY == variety.temp,3] <- rYEARS.temp[1]
  varieties[varieties$VARIETY == variety.temp,4] <- rYEARS.temp[2]
  varieties[varieties$VARIETY == variety.temp,5] <- nrow(variety.years.temp)
}
names(varieties) <- c("VARIETY", "GAPS", "START", "END", "NUM.EXPERIMENTS")

num.varieties <- nrow(varieties)

num.gaps <- sum(varieties$GAPS)



```

**There is a total of `r num.gaps` varieties with gaps in its experimentation process.**

`r varieties[varieties$GAPS == 1,] `

#### <span style="color:green"> Following 5 different varieties </span>

To discover interesting varieties, for each variety, counted number of distinct years of experiments.
<BR>
<BR>

```{r, echo= FALSE, fig.width= 10, fig.align='center'}

p <- plot_ly(data = varieties, x = NUM.EXPERIMENTS, color = GAPS, type = "histogram")

x.axis <- list(title = "Number of Years of Experimentation")
y.axis <- list(title = "Count of varieties")
p %>% layout(title = paste("Number of years of experimentation. Count:", num.varieties, "varieties."),
             xaxis = x.axis, yaxis = y.axis)


```

<br>

I will pick one variety from each stack from 2 to 6.

```{r, echo= FALSE}

varieties <- varieties[order(varieties$VARIETY, decreasing = TRUE),]

variety2 <- varieties[varieties$NUM.EXPERIMENTS == 2,]
variety2 <- as.character(variety2[1,1])

variety3 <- varieties[varieties$NUM.EXPERIMENTS == 3,]
variety3 <- as.character(variety3[1,1])

variety4 <- varieties[varieties$NUM.EXPERIMENTS == 4,]
variety4 <- as.character(variety4[1,1])

variety5 <- varieties[varieties$NUM.EXPERIMENTS == 5,]
variety5 <- as.character(variety5[1,1])

variety6 <- varieties[varieties$NUM.EXPERIMENTS == 6,]
variety6 <- as.character(variety6[1,1])

# Plotting function
plot.histogram <- function(experiment.data.input){

  
  experiment.data.aggr <- unique(experiment.data.input[c("YEAR", "EXPERIMENT", "LOCATION", "VARIETY", "CHECK")])
    p <- plot_ly(data = experiment.data.aggr, x = YEAR, color = CHECK, type = "histogram", 
                 nbinsx = max(experiment.data.aggr$YEAR) - min(experiment.data.aggr$YEAR) + 1)
  x.axis <- list(title = "Year")
  y.axis <- list(title = "Count of (Experiments,Location)")
  p %>% layout(title = paste("Number of (Experiments,Location). Count:", nrow(experiment.data.aggr)),
               xaxis = x.axis, yaxis = y.axis)
}


```

The selected varieties are the following:

Number of Years | Variety
--------------- | -----------
2               | `r variety2`
3               | `r variety3`
4               | `r variety4`
5               | `r variety5`
6               | `r variety6`

<BR>

### Variety with 2 years: `r variety2`
<BR>

```{r, echo= FALSE}
experiment.data.2 <- experiment.data[experiment.data$VARIETY == variety2,]
datatable(experiment.data.2)
```
<BR>
```{r, echo= FALSE, fig.align='center'}
plot.histogram(experiment.data.2)
```



### Variety with 3 years: `r variety3` (This seed has a gap in 2012!)
<BR>
```{r, echo= FALSE}
experiment.data.3 <- experiment.data[experiment.data$VARIETY == variety3,]
datatable(experiment.data.3)
```

```{r, echo= FALSE, fig.align='center'}
plot.histogram(experiment.data.3)
```

<BR>

### Variety with 4 years: `r variety4`
<BR>
```{r, echo= FALSE}
experiment.data.4 <- experiment.data[experiment.data$VARIETY == variety4,]
datatable(experiment.data.4)
```

```{r, echo= FALSE, fig.align='center'}
plot.histogram(experiment.data.4)
```
<BR>

### Variety with 5 years: `r variety5`
<BR>
```{r, echo= FALSE}
experiment.data.5 <- experiment.data[experiment.data$VARIETY == variety5,]
datatable(experiment.data.5)
```
```{r, echo= FALSE, fig.align='center'}
plot.histogram(experiment.data.5)
```

<BR>

### Variety with 6 years: `r variety6`
<BR>

```{r, echo= FALSE}
experiment.data.6 <- experiment.data[experiment.data$VARIETY == variety6,]
datatable(experiment.data.6)
```

```{r, echo= FALSE, fig.align='center'}
plot.histogram(experiment.data.6)
```



