---
title: "Experiment Tree Analysis"
author: "Sin City Seeders"
date: "October 17, 2016"
output: 
  html_notebook: default
---


```{r setup, include=FALSE}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(out.width='750px', dpi=200)

```

```{r, include=FALSE}
library(dplyr)
library(DT)
library(readr)
library(plotly)
library(reshape2)
library(readxl)
```


```{r, include=FALSE}
experiment.data.tree <- read_csv("../data/exp_data_final.csv", 
                                 col_types = cols(
                                   LOCATION = col_character()
                                 ))
```


<br>
<br>
<br>

<H3><CENTER>High Level Table</H3></CENTER>
Understanding how many unique varieties are present in each stage of each tree

**Remember that 80%+ of varieties are discarded in every stage.**
<br>
<br>

<CENTER>Rows: STARTING_YEAR, Columns: STAGE</CENTER>


```{r, echo=FALSE}
uniques.count <- group_by(experiment.data.tree, STARTING_YEAR, STAGE)
uniques.count <- summarise(uniques.count, count.varieties = n_distinct(VARIETY))
uniques.count.transpose <- dcast(uniques.count, STARTING_YEAR ~ STAGE, value.var = "count.varieties")

total.varieties.stage.1 <- sum(uniques.count.transpose$`1`)
total.varieties.not.check <- nrow(unique(experiment.data.tree["VARIETY"]))
datatable(uniques.count.transpose)



plot.tree.evolution.histogram <- function(starting.year){
  experiment.data.tree.year <- experiment.data.tree[experiment.data.tree$STARTING_YEAR == starting.year,]
  experiment.data.tree.year.uniques <- unique(experiment.data.tree.year[c("YEAR", "VARIETY", "QUALIFIED", "CHECK")])
  
  total.varieties <- nrow(experiment.data.tree.year.uniques[experiment.data.tree.year.uniques$YEAR == starting.year,])
  
  
  p <- plot_ly(data=experiment.data.tree.year.uniques, x = YEAR, color = QUALIFIED, type = "histogram")
  
  x.axis <- list(title = "Year")
  y.axis <- list(title = "Count of Varieties")
  
  p <- p %>% layout(title = paste0("Count of Unique Varieties Tested per year. <br> Starting year: ",starting.year,
                             ". Total Varieties: ", total.varieties),
               xaxis = x.axis, 
               yaxis = y.axis,
               barmode = "stack")
  
  dt <- datatable(experiment.data.tree.year.uniques)

  htmltools::tagList(list(as.widget(p), as.widget(dt)))
  
}

```

<br>
<br>

<UL>
<LI> Very few varieties qualify in each year! For example, for 2009 Stage 1, 71 / 1789 = 4% qualified.</LI> 
<LI> This datatable has a total of `r total.varieties.stage.1` unique varieties in stage 1. </LI>
<LI> There is a total of `r total.varieties.not.check` unique varieties in the exp_data_final database. </LI> 
<LI> There could be `r (total.varieties.stage.1 - total.varieties.not.check)` varieties duplicated. </LI>
<LI> This represents `r round((total.varieties.stage.1 - total.varieties.not.check) / total.varieties.not.check, digits = 4)*100`% of the varieties. </LI>
</UL>


<br>
<br>

***
<H3><CENTER>Histograms for each STARTING_YEAR </CENTER></H3>
***

```{r, echo=FALSE, fig.align='center', warning=FALSE}
plot.tree.evolution.histogram(2009)
plot.tree.evolution.histogram(2010)
plot.tree.evolution.histogram(2011)
plot.tree.evolution.histogram(2012)
plot.tree.evolution.histogram(2013)
plot.tree.evolution.histogram(2014)
```


<H3><CENTER>Discussion, Further Work</H3></CENTER>


* **[POSITIVE] The bulk of the data seems to follow INFORMS' story.**
    + In all trees, seeds are filtered every year, and very few qualify.
* [FURTHER WORK] Have the full experiment database, with this 3 new columns. 
    + This is a subset of the entire database.
    + Elements used as check maybe just keep these 3 columns blank.
* [FURTHER WORK] With these new columns, can we now explain the missing values of CLASS_OF and GRAD?  
* [FURTHER WORK] With these new columns, can we now explain the missing values of CLASS_OF and GRAD?
* [FURTHER WORK] In 2012, 2014 we have some varieties with no data in QUALIFIED
    + In 2012 means those are the seeds WE have to make the decision?
    + Contrast Varieties present in START_YEAR = 2012, STAGE = 3, with the list of varieties with the EVALUATION SET.
* [FURTHER WORK] Why are we even testing in stage 1 of 2013, 2014 ? 



```{r, echo=FALSE, fig.align='center', warning=FALSE}
full.dataset <-  read_excel(path = '../data/EXPERIMENT DATA.xlsx', 
                              sheet = "TRAINING_DATA", col_names = TRUE)


relevant.columns <- names(full.dataset)
relevant.columns <- relevant.columns[relevant.columns != "YIELD"]
full.dataset.relevant <- full.dataset[,relevant.columns]
experiment.data.tree.relevant <- experiment.data.tree[,relevant.columns]

```


